Depression and suicidality as costly signals of need

### 2021-04-08 Paper.rmd
Worked on the discussion - Added discussion about why suicide attempt may have resulted in greater likelihood of action by not belief. This involved merging the main effect and mediation sections in addition to other edits.

## notes: 
add conclusion
I cannot knit (works with library targets in chunk starting at line 13). But then i get an error bc I don't have hrafsuicidedata package

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable increase in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second increase when going from depression to a suicide attempt with the "Action' variable. One potential reason for former is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lengthy period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need.

## Mediation models

Our prediction that the effect of signal cost on likelihood of help would largely be mediated by signal's effect on belief in need was also met, as would be expected if the main function of signaling need is providing trustworthy information about one's internal state to others [@Hagen2003]. However, there was still a positive main effect of the signal in many pairwise comparisons between the costlier signals and the less costly signals, including the comparison between the depression and verbal request condition (See SI for other mediation models). One possible reason for this is that someone may still be distressed and face real adversity without being forthcoming about one's problems. For example, participants who saw the "brother-in-law" vignette may have thought the daughter was lying about the assault due to jealously over her cousin, but the stronger signals may have still conveyed the severity of the competition between her and her cousin in ways words alone could not.

@@

Our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. In addition to the effect of signal cost on support being largely mediated by the signal's effect on belief in need, belief in need and likelihood of support were highly correlated but distinct at T1 (pre-signal). This situation should arise in any instance when a target of a signal of need would benefit from helping if the need was genuine and known to reasonable precision. Importantly, such problems are common among modern humans and have likely been so since individuals could increase their inclusive fitness through supporting kin or important cooperative partners in times of genuine need [@SugiyamaSugiyama2003; @tooby1996].

@@

The main limitation of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction if the lack of meaningful costs made support feel less costly or if the likely inability of participants to imagine themselves in an anonymous scenario [@hagen2006] led them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the costlier signals as bargaining tools. However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs.

@@

>>>>

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information). It is also possible that our research in designing the vignettes was insufficient for the population or the content of the vignette.

In the "brother-in-law" and "romantic partner" scenarios, *Crying* had little effect on the magnitude of pro-signaler responses, while both *Depression* conditions increased support to similar degrees. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lengthy period of time.

While the effect of *Suicide attempt* on belief at T2 was similar to the *Depression* conditions, it resulted in greater likelihood of support. This may be due to our decision to measure belief in need as the degree participants believed the claims of the signaler (i.e., the assaults or incompatibility with a potential husband in the context of arranged marriage) rather than a more inclusive measure. For this reason, signal cost may have provided additional information about the costliness of the signaler's adversity in ways that would still be consistent with signaling theory.

Support for this interpretation comes from our mediation analyses, which showed that the likelihood of help was largely, but not entirely, mediated by signal's effect on belief in need in this study but totally mediated by belief in need when measured as belief in need of financial support in the pilot study. However, the incomplete mediation in this study may have arisen if participants viewed the signal as providing information about adveristy outside of the signaler's claims. For example, participants who saw the "brother-in-law" vignette may have thought the daughter was lying about the assault due to jealously over her cousin, with the stronger signals still conveying the severity of the competition between her and her cousin.

@@

Our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. In addition to the effect of signal cost on support being largely mediated by the signal's effect on belief in need, belief in need and likelihood of support were highly correlated but distinct at T1. This situation should arise in any instance when a target of a signal of need would benefit from helping if the need was genuine and known to reasonable precision. Importantly, such problems are common among modern humans and have likely been so since individuals could increase their inclusive fitness through supporting kin or important cooperative partners in times of genuine need [@SugiyamaSugiyama2003; @tooby1996].

@@

The main limitation of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction if the lack of meaningful costs made support feel less costly or if the likely inability of participants to imagine themselves in an anonymous scenario [@hagen2006] led them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the costlier signals as bargaining tools. However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity (needs improvement. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs.

@@

# To do:
- add vignettes to SI
- mention somewhere referring to them all as signals even though we know they provide info as cues too
